---
title: "Getting familiar with dMrs"
author: "Paul Little, Reuben Adatorwovor"
date: "Last Updated: `r format(Sys.Date(), '%m/%d/%Y')`"
output: 
  html_document:
    theme: journal
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Getting familiar with ProjectH}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r global_options,include = FALSE}
knitr::opts_chunk$set(comment = NA)
```

# Introduction
This **dMrs** package is designed to fit survival data to the corresponding manuscript.

```{r load_lib,echo = TRUE,eval = TRUE}
library(dMrs)
```

# Simulation

Specify underlying arguments and parameters.

```{r set_parameters,echo = TRUE,eval = TRUE}
copula = c("Clayton","Gumbel")[1]
dist1 = c("weibull","expweibull")[1]
NN = 2e2 # sample size
theta = 3/2
alpha1 = 5
lambda1 = 1.2
kappa1 = 2
kappa1 = ifelse(dist1 == "weibull",1,kappa1)
alpha2 = 2
lambda2 = 2.1
propC = 0.1
```

## One replicate

Simulate a dataset.

```{r gen_data,echo = TRUE,eval = TRUE}
set.seed(1)
one_rep = sim_replicate(copula = copula,dist1 = dist1,
  NN = NN,theta = theta,alpha1 = alpha1,lambda1 = lambda1,
  kappa1 = kappa1,alpha2 = alpha2,lambda2 = lambda2,
  propC = propC,show = TRUE)
str(one_rep)
head(one_rep$DATA)
table(one_rep$DATA$D)
table(one_rep$DATA$delta)
```

For a given parameter grid and theta value, first, run a grid search. 
If `show = TRUE`, verbose output is provided and a heatmap of the log likelihood is generated.
Then run parameter optimization for the true `theta` value and an arbitrary `theta` value for fixed `theta` (`upTHETA = 0`). We are also assuming the first event time is Weibull distributed (`upKAPPA = 0`).

```{r opt_sim,echo = TRUE,eval = TRUE,fig.dim = c(12,7)}
opt_out = opt_replicate(REP = one_rep,
  param_grid = seq(-1,3,0.25),
	theta = one_rep$PARAMS$theta,
  upKAPPA = 0,
  upTHETA = 0,
  show = TRUE)
str(opt_out)

opt_out = opt_replicate(REP = one_rep,
  param_grid = seq(-1,3,0.25),
  theta = 2,
  upKAPPA = 0,
  upTHETA = 0,
  show = TRUE)
str(opt_out)
```

We can determine if the optimization converged or not by looking at the gradient and check out other quantities when evaluated at the MLE.

```{r}
# MLEs (unconstrained)
opt_out$out

# MLEs (constrained)
opt_out$cout

# Log-likelihood
opt_out$LL

# Gradient, Score, first derivative of log-likelihood
opt_out$GRAD

# Hessian, second derivative of log-likelihood
opt_out$HESS

# Covariance matrix
opt_out$COVAR

```

Want to re-generate the heatmap of the log likelihood again?

```{r}
plot_LL(GPROF = opt_out$GPROF,GOPT = opt_out$GOPT)
```

## Full simulation, multiple replicates

The code below performs a full simulation across multiple replicates.

```{r full_sim,echo = TRUE,eval = TRUE,fig.dim = c(7,4)}
RR = 1e2 # number of replicates

sim_out = full_sim(copula = copula,
  dist1 = dist1,NN = NN,theta = theta,
  alpha1 = alpha1,lambda1 = lambda1,
  kappa1 = kappa1,alpha2 = alpha2,
  lambda2 = lambda2,propC = propC,RR = RR,
  param_grid = seq(-1,3,0.5),
  upKAPPA = 1,show = TRUE)
sim_out
```

# Applying dMrs

Equivalently, we can run a formal analysis for various thetas, again with the simulated dataset. 

```{r ana,echo = TRUE,eval = TRUE,fig.dim = c(12,7)}
ana_out = run_analyses(DATA = one_rep$DATA,
	THETAs = c(0,one_rep$PARAMS$theta,2,5),copula = copula,
	param_grid = seq(-1,3,0.1),
	vec_time = seq(0,2 * max(one_rep$DATA$time),length.out = 50),
  upKAPPA = 0,show = TRUE)
names(ana_out)

# Estimates per theta
bb = sapply(names(ana_out),function(xx){
	# xx = names(run_ana)[1]; xx
	cat("\n#######\n")
	print(xx)
	print(ana_out[[xx]]$RES$cout[,c("PARS","EST","lowCI_2","highCI_2")])
	print(sprintf("LL = %s",round(ana_out[[xx]]$RES$LL,3)))
	print(round(ana_out[[xx]]$PRED[1:5,c("time","surv",
		"low_surv2","high_surv2")],3))
})

```

Below we plot the corresponding predicted survival with 95% confidence bands. In the latter figure, we can see the corresponding maximum log likelihood estimates per theta.

```{r plot_surv,echo = TRUE,eval = TRUE,fig.dim = c(12,7)}
plot_SURVs(run_ANA = ana_out,MULTIPLE = FALSE,ALPHA = 0.5)
plot_SURVs(run_ANA = ana_out,MULTIPLE = TRUE,ALPHA = 0.5)
```

# Session information

```{r session,echo = FALSE}
sessionInfo()
```

# References

